### 2025-06-29 machine-learning.qmd
20250629 - notes to self

CURRENTLY (WILL CHANGE):

- seasonalData: 1999-2024: i.e. all data currently available
- seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
- newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)

to create:

- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players; 2023 rather than 2024 because don't have fantasy points for retired players in most recent year
- (eventually, newData_seasonal, which is derived from the imputed version of seasonalData_lag_all, and then keeps only data from 2024, thus predicting fantasy points in 2025, and removing the fantasy_points_lag variable)

To impute:

- seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
- seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most players
- seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some of the) retired players

IMPUTATION AND MODEL PROCESS:

IMPUTATION:
- training data
- test data
- all data (training and test data), used for generating next year predictions

MODEL:
- training data (imputed version of seasonalData_lag_all): all seasons (except 2024 predicting 2025), most players

EVALUATE MODEL PREDICTIONS:
- test data (imputed version of seasonalData_lag_test): all seasons (except 2024 predicting 2025), some retired players

GENERATE MODEL PREDICTIONS
- next year data (newData_seasonal): 2024 (predicting 2025)

@@

NOTES:

- CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
- CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
- CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION

@@

data_test_qb <- seasonalData_lag_qb_test_imp
data_test_qb_matrix <- data_test_qb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

# RBs
seasonalData_lag_rb_all_imp <- missRanger::missRanger(
  seasonalData_lag_rb_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

@@

data_test_rb <- seasonalData_lag_rb_test_imp
data_test_rb_matrix <- data_test_rb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

# WRs
seasonalData_lag_wr_all_imp <- missRanger::missRanger(
  seasonalData_lag_wr_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

@@

data_test_wr <- seasonalData_lag_wr_test_imp
data_test_wr_matrix <- data_test_wr %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()

# TEs
seasonalData_lag_te_all_imp <- missRanger::missRanger(
  seasonalData_lag_te_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

>>>>

# Notes to self

# CURRENTLY (WILL CHANGE):
# 
# - seasonalData: 1999-2024: i.e. all data currently available
# - seasonalData_lag: 1999-2023 (predicting fantasy points in 2024)
# - newData_seasonal: 2024 (to be used for predicting fantasy points in 2025)

# to create:
# 
# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players; 2023 rather than 2024 because don't have fantasy #points for retired players in most recent year
# - (eventually, newData_seasonal, which is derived from the imputed version of #seasonalData_lag_all, and then keeps only data from 2024, thus predicting #fantasy points in 2025, and removing the fantasy_points_lag variable)

# To impute:
# 
# - seasonalData_lag_all: 1999-2023 (predicting fantasy points in 2024)
# - seasonalData_lag_train: 1999-2023 (predicting fantasy points in 2024), most #players
# - seasonalData_lag_test: 1999-2022 (predicting fantasy points in 2023), (some #of the) retired players

# IMPUTATION AND MODEL PROCESS:
# 
# IMPUTATION:
# - training data
# - test data
# - all data (training and test data), used for generating next year #predictions
# 
# MODEL:
# - training data (imputed version of seasonalData_lag_all): all seasons #(except 2024 predicting 2025), most players
# 
# EVALUATE MODEL PREDICTIONS:
# - test data (imputed version of seasonalData_lag_test): all seasons (except #2024 predicting 2025), some retired players
# 
# GENERATE MODEL PREDICTIONS
# - next year data (newData_seasonal): 2024 (predicting 2025)
```

@@

# Notes to self

# - CONSIDER USING RETIRED PLAYERS AS THE HOLDOUT SAMPLE
# - CONSIDER WIDENING THE DATA TO AVOID MULTILEVEL IMPUTATION
# - CONSIDER IMPUTING THE TRAINING AND TEST DATA SEPARATELY BY POSITION
```

@@

data_test_qb <- seasonalData_lag_qb_test_imp
data_test_qb_matrix <- data_test_qb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
# RBs
seasonalData_lag_rb_all_imp <- missRanger::missRanger(
  seasonalData_lag_rb_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

@@

data_test_rb <- seasonalData_lag_rb_test_imp
data_test_rb_matrix <- data_test_rb %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
# WRs
seasonalData_lag_wr_all_imp <- missRanger::missRanger(
  seasonalData_lag_wr_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)

@@

data_test_wr <- seasonalData_lag_wr_test_imp
data_test_wr_matrix <- data_test_wr %>%
  mutate(across(where(is.factor), ~ as.numeric(as.integer(.)))) %>% 
  as.matrix()
```

```{r}
# TEs
seasonalData_lag_te_all_imp <- missRanger::missRanger(
  seasonalData_lag_te_all,
  pmm.k = 5,
  verbose = 2,
  seed = 52242,
  keep_forests = TRUE)
